---
title: "Short Analysis of Lubridillo Site"
format: html
editor: visual
execute:
  warning: false
  message: false
---

# Loading Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggExtra)
library(lubridate)
library(patchwork)
```

# Documentation

Group:

1\) Created shared folder on Taylor-Adelaide

2\) Created github repository and shared with collaborators-Ale

3)Created Quarto graph with sections to work in separately

Andre - work on graph

# Import, tidy, and join data

```{r}

# read all quebradas
setwd("/courses/EDS214/pr/data/knb_lter_luq_raw")

p_r_m <- read_csv("RioMameyesPuenteRoto.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

bq_1 <- read_csv("QuebradaCuenca1-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
bq_2 <- read_csv("QuebradaCuenca2-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
bq_3 <- read_csv("QuebradaCuenca3-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))
  
# merge data via full join 
stream_water_df <- p_r_m %>% 
 full_join(bq_1) %>% 
  full_join(bq_2) %>%
  full_join(bq_3)

# export the data frame
setwd("/courses/EDS214/pr/data/processed_data")
write_csv(stream_water_df, "stream_water.csv")
```

# Recreate Graph (Step 1)

## WORK UNDER HERE

```{r}

stream_water_clean <- stream_water_df %>%
  select(sample_id, sample_date_ymd, year, month, day, k, no3_n) %>%
  filter(year(sample_date_ymd) %in% 1989:1992)

start_date <- ymd("1989-05-16")
end_date <- ymd("1992-04-30")

date <- interval(start_date, end_date)

prm <- stream_water_clean %>%
  select(sample_id, sample_date_ymd, month, k, no3_n) %>%
  filter(sample_id == "MPR" & year(sample_date_ymd) %in% 1989:1992) 

bq_1 <- stream_water_clean %>%
  select(sample_id, sample_date_ymd, month, k, no3_n) %>%
  filter(sample_id == "Q1" & year(sample_date_ymd) %in% 1989:1992) 

bq_2 <- stream_water_clean %>%
  select(sample_id, sample_date_ymd, month, k, no3_n) %>%
  filter(sample_id == "Q2" & year(sample_date_ymd) %in% 1989:1992) 

bq_3 <- stream_water_clean %>%
  select(sample_id, sample_date_ymd, month, k, no3_n) %>%
  filter(sample_id == "Q3" & year(sample_date_ymd) %in% 1989:1992) 

unique(stream_water_clean$sample_id)

```

```{r}
joined_data <- bind_rows(bq_1, bq_2, bq_3, prm)
## select to keep rows: sample_id, sample_date, no3_n, k
stream_quality <- joined_data %>%
  select(sample_id, sample_date_ymd, no3_n, k) %>%
  mutate(year = year(sample_date_ymd),
         date = ymd(paste0(year, "-", month(sample_date_ymd), "-1"))) %>%
  filter(year %in% 1988:1994)
monthly_quality <- stream_quality %>%
  group_by(sample_id, date) %>%
  summarize(monthlyno3 = mean(no3_n, na.rm = TRUE),
            monthlyk = mean(k, na.rm = TRUE))
```

```{r}
ggplot(monthly_quality, aes(x = date,
                               y = monthlyk) )+
         geom_line(aes(color = sample_id))


```

## WORK UNDER HERE ALSO

```{r}
ggplot(monthly_quality, aes(x = date,
                               y = no3_n)) +
  geom_line(aes(color = sample_id))
```

# Suspended sediment and temperature

```{r}
sediment_temp_df <- stream_water_df %>%
  select(-c(gage_ht, p_h, cond, cl, no3_n,
            so4_s, na, k, mg, ca, nh4_n, po4_p,
            tdn, tdp, si_o2, don, dic, doc,
            chloride_code, nitrate_code, 
            sulfate_code, sodium_code, 
            potassium_code, magnesium_code, calcium_code,
            nh4code, po4code, doc_code,
            dic_code, tdn_code, si_o2code,
            don_code)) %>%
  mutate(tss = as.numeric(tss))

sediment_temp_long <- sediment_temp_df %>%
  pivot_longer(temp, tss, names_to = "parameter")
```

```{r}
ggplot(sediment_temp_df) +
  geom_line(aes(x = sample_date_ymd, 
                y = temp),
            color = "red") + 
  geom_line(aes(x = sample_date_ymd,
                y = tss),
            color = "blue")
```

Over time how to total sediment and carbon vary and is there a
correlation between the two.

<https://www.nature.com/articles/s41598-018-33640-3>

https://link.springer.com/article/10.1007/s40996-020-00460-3

# Different Method

```{r}

sediment_temp_df <- stream_water_df %>%
  select(-c(gage_ht, p_h, cond, cl, no3_n,
            so4_s, na, k, mg, ca, nh4_n, po4_p,
            tdn, tdp, si_o2, don, dic, doc,
            chloride_code, nitrate_code, 
            sulfate_code, sodium_code, 
            potassium_code, magnesium_code, calcium_code,
            nh4code, po4code, doc_code,
            dic_code, tdn_code, si_o2code,
            don_code)) %>%
  mutate(tss = as.numeric(tss)) %>% mutate(year = year(sample_date)) 
#made a column with year

sediment_temp_long <- sediment_temp_df %>%
  pivot_longer(cols = c(temp, tss), names_to = "parameter") #put into long format
%>%
  group_by(year,parameter) %>% 
  summarize( mean_value = mean(value, na.rm = TRUE)) # take average of temp and tss
```

```{r}

ggplot(sediment_temp_long, aes(x = year, y = mean_value)) + geom_line() +
  facet_wrap(~parameter, scales = "free", ncol = 1)
```
