---
title: "Short Analysis of Lubridillo Site"
format: html
editor: visual
execute:
  warning: false
  message: false
---

# Loading Libraries

```{r}
library(tidyverse)
library(here)
library(janitor)
library(ggExtra)
library(lubridate)
library(patchwork)
library(RCurl)
```

# Division of tasks

-   Created shared folder on Taylor-Adelaide

-   Created github repository and shared with collaborators-Alessandra

-   Created Quarto document with sections to work in separately- Group Effort

-   Import and tidy data- Group Effort

-   Graph Recreation-Andre/Guillermo

-   New Question- Alessandra/Adelaide

-   Slides- Guillermo

# Import, tidy, and join data

```{r}

# Update current directory (cd):
setwd("/courses/EDS214/pr/data/knb_lter_luq_raw")

prm <- read_csv("RioMameyesPuenteRoto.csv", na = c("NA")) %>%   
  clean_names() %>% #tidy the data to snake_case
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>% # Adjust to ymd format
  mutate(year = year(sample_date_ymd), 
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

bq_1 <- read_csv("QuebradaCuenca1-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
bq_2 <- read_csv("QuebradaCuenca2-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
bq_3 <- read_csv("QuebradaCuenca3-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))
  
# merge data via full join 
stream_water_df <- prm %>% 
 full_join(bq_1) %>% 
  full_join(bq_2) %>%
  full_join(bq_3)

# export the data frame
setwd("/courses/EDS214/pr/data/processed_data")
write_csv(stream_water_df, "stream_water.csv")
```

# Recreate Graph

```{r}
stream_quality <- stream_water_df %>%
   select(sample_id, sample_date_ymd, no3_n, k) %>% #K and No3N data by watershed (sample_id)
  mutate(year = year(sample_date_ymd),
         date = ymd(paste0(year, "-", month(sample_date_ymd), "-1"))) %>%
  filter(year %in% 1988:1994) #Year data range

monthly_stream_quality <- stream_quality %>%
  group_by(sample_id, date) %>% #grouped data by date and watershed
  summarize(monthlyno3 = mean(no3_n, na.rm = TRUE), #Get the monthly mean
            monthlyk = mean(k, na.rm = TRUE)) %>%
  mutate(sample_id = str_replace(sample_id, "MPR", "Puente Roto Mameyes")) %>%
  mutate(sample_id = str_replace(sample_id, "Q1", "Quebrada 1 - Bisley")) %>%
  mutate(sample_id = str_replace(sample_id, "Q2", "Quebrada 2 - Bisley")) %>%
  mutate(sample_id = str_replace(sample_id, "Q3", "Quebrada 3 - Bisley"))
```

```{r}
ggplot(monthly_stream_quality, aes(x = date,
                               y = monthlyk) )+
         geom_line(aes(color = sample_id)) +
  labs(title = "Potassium Levels by Watershed",
       caption = "McDowell, W. 2022. Chemistry of stream water \nfrom the Luquillo Mountains ver 4923058. Environmental Data Initiative.\nhttps://doi.org/10.6073/pasta/1dd1a59a7e6fbe95fdf8736477d81b83",
       x = "Year",
       y = "Potassium Levels (mg/L)") +
  facet_wrap(~sample_id) + #facet by different watersheds
  theme_test() + 
  theme(legend.position = "none", #facet, legend not necessary
        panel.grid.major = element_line(colour = "gray",
                                        size = 0.10),
        plot.title.position = "plot") #Shift title to the left
  
 
```

```{r}
ggplot(monthly_stream_quality, aes(x = date,
                               y = monthlyno3) )+
         geom_line(aes(color = sample_id)) +
  labs(title = "Nitrate Nitrogen Levels by Watershed",
       caption = "McDowell, W. 2022. Chemistry of stream water \nfrom the Luquillo Mountains ver 4923058. Environmental Data Initiative.\nhttps://doi.org/10.6073/pasta/1dd1a59a7e6fbe95fdf8736477d81b83",
       x = "Year",
       y = "Nitrate Nitrogen Levels (mg/L)") +
  facet_wrap(~sample_id) +
  theme_test() +
  theme(legend.position = "none",
        panel.grid.major = element_line(colour = "gray",
                                        size = 0.10),
        plot.title.position = "plot")
```

# 

# New Question

Looking at the relationship between total suspended sediment and pH for the same three sites from 1989 to 2017.

```{r}
sediment_ph_df <- stream_water_df %>%
  select(-c(gage_ht, temp, cond, cl, no3_n,
            so4_s, na, k, mg, ca, nh4_n, po4_p,
            tdn, tdp, si_o2, don, dic, doc,
            chloride_code, nitrate_code, 
            sulfate_code, sodium_code, 
            potassium_code, magnesium_code, calcium_code,
            nh4code, po4code, doc_code,
            dic_code, tdn_code, si_o2code,
            don_code)) %>% # selecting relevant columns
  mutate(tss = as.numeric(tss)) %>% # changing tss to numeric
  mutate(yearc = as.character(year)) #changing to character for linear regression fixed effects 

sediment_ph_long <- sediment_ph_df %>%
  pivot_longer(cols = c(p_h, tss), names_to = "parameter") %>% #combining p_h and tss into one column
  group_by(year, parameter, sample_id) %>% 
  summarize(mean_parameter = mean(value, na.rm = TRUE)) %>% # find mean of tss and ph by year
  mutate(sample_id = str_replace(sample_id, "MPR", "Puente Roto Mameyes"),
         sample_id = str_replace(sample_id, "Q1", "Quebrada 1 - Bisley"),
         sample_id = str_replace(sample_id, "Q2", "Quebrada 2 - Bisley"),
         sample_id = str_replace(sample_id, "Q3", "Quebrada 3 - Bisley"),
         parameter = str_replace(parameter, "p_h", "PH"),
         parameter = str_replace(parameter, "tss", "Total Suspended Sediment (mg/L)")) # relabeling sites and parameters
```

```{r}
ggplot(sediment_ph_long,
       aes(x = year,
           y = mean_parameter)) +
  geom_line(aes(color = sample_id)) + #split each site into a line
  facet_wrap(~parameter,
             scales = "free",
             ncol = 1) + #split graphs by paramater
  theme(legend.title = element_blank(),
        legend.position = "bottom") +
  geom_vline(xintercept = 1989, 
             linetype = "dashed") + 
  geom_vline(xintercept = 2017, 
             linetype = "dashed") +  #adding vertical lines to indicate hurricanes
  labs(title = "PH and total suspended sediment in stream water \nfrom the Luquillo Mountains",
       caption = "McDowell, W. 2022. Chemistry of stream water \nfrom the Luquillo Mountains ver 4923058. Environmental Data Initiative.\nhttps://doi.org/10.6073/pasta/1dd1a59a7e6fbe95fdf8736477d81b83",
       x = "Year",
       y = NULL)
```

```{r}
# create subsets of summary_ph_df data
stream_water_df_1988 <- subset(sediment_ph_df, year == 1988)
stream_water_df_1989 <- subset(sediment_ph_df, year == 1989)
stream_water_df_1990 <- subset(sediment_ph_df, year == 1990)
```

```{r}
# Linear regression for 1988
sediment_ph_1988_lm <- lm (p_h ~ tss + sample_id, stream_water_df_1988)
summary(sediment_ph_1988_lm, robust = T)

# Linear regression for 1989
sediment_ph_1989_lm <- lm (p_h ~ tss + sample_id, stream_water_df_1989)
summary(sediment_ph_1989_lm, robust = T)

# First difference
f_diff_1989_88 <- lm(I(stream_water_df_1989$p_h - stream_water_df_1988$p_h)~I(stream_water_df_1989$tss - stream_water_df_1988$tss))
summary(f_diff_1989_88)

f_diff_1990_88 <- lm(I(stream_water_df_1990$p_h - stream_water_df_1988$p_h)~I(stream_water_df_1990$tss - stream_water_df_1988$tss))
summary(f_diff_1990_88)

```

```{r}
ggplot(stream_water_df_1988,
       aes(x = p_h,
           y = tss)) + 
  geom_point(color = "#32965D",
             alpha = 1/4,
             shape = 19) +
  stat_smooth(method = "lm", col = "#68A357") +
  theme_light() +
  labs(title = "Linear regression of PH and total suspended sediment \nfrom the Luquillo Mountains in 1988",
       caption = "McDowell, W. 2022. Chemistry of stream water \nfrom the Luquillo Mountains ver 4923058. Environmental Data Initiative.\nhttps://doi.org/10.6073/pasta/1dd1a59a7e6fbe95fdf8736477d81b83",
       x = "PH",
       y = "Total Suspended Sediment (mg/L)")
```

```{r}
ggplot(stream_water_df_1990,
       aes(x = p_h,
           y = tss)) + 
  geom_point(color = "#32965D",
             alpha = 1/4,
             shape = 19) +
  stat_smooth(method = "lm", col = "#68A357") +
  theme_light() +
  labs(title = "Linear regression of PH and total suspended sediment \nfrom the Luquillo Mountains in 1990",
       caption = "McDowell, W. 2022. Chemistry of stream water \nfrom the Luquillo Mountains ver 4923058. Environmental Data Initiative.\nhttps://doi.org/10.6073/pasta/1dd1a59a7e6fbe95fdf8736477d81b83",
       x = "PH",
       y = "Total Suspended Sediment (mg/L)")
```
