---
title: "Short Analysis of Lubridillo Site"
format: html
editor: visual
execute:
  warning: false
  message: false
---

# Loading Libraries

```{r}
library(tidyverse)
library(janitor)
library(lubridate)
```

# Documentation

Group:

1\) Created shared folder on Taylor-Adelaide

2\) Created github repository and shared with collaborators-Ale

3)Created Quarto graph with sections to work in separately

Andre - work on graph

# Import, tidy, and join data

```{r}

# read all quebradas
setwd("/courses/EDS214/pr/data/knb_lter_luq_raw")

p_r_m <- read_csv("RioMameyesPuenteRoto.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

b_q_1 <- read_csv("QuebradaCuenca1-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
b_q_2 <- read_csv("QuebradaCuenca2-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))

  
b_q_3 <- read_csv("QuebradaCuenca3-Bisley.csv", na = c("NA")) %>%
  clean_names() %>%
  mutate(sample_date_ymd = lubridate::ymd(sample_date)) %>%
  mutate(year = year(sample_date_ymd),
         month = month(sample_date_ymd),
         day = day(sample_date_ymd))
  
# merge data via full join 
stream_water_df <- p_r_m %>% 
  full_join(b_q_1) %>% 
  full_join(b_q_2) %>%
  full_join(b_q_3)

# export the data frame
setwd("/courses/EDS214/pr/data/processed_data")
write_csv(stream_water_df, "stream_water.csv")
```

# Recreate Graph (Step 1)

## WORK UNDER HERE

```{r}

```

## WORK UNDER HERE ALSO

```{r}

```

# Suspended sediment and temperature

```{r}
sediment_temp_df <- stream_water_df %>%
  select(-c(gage_ht, p_h, cond, cl, no3_n,
            so4_s, na, k, mg, ca, nh4_n, po4_p,
            tdn, tdp, si_o2, don, dic, doc,
            chloride_code, nitrate_code, 
            sulfate_code, sodium_code, 
            potassium_code, magnesium_code, calcium_code,
            nh4code, po4code, doc_code,
            dic_code, tdn_code, si_o2code,
            don_code)) %>%
  mutate(tss = as.numeric(tss))

sediment_temp_long <- sediment_temp_df %>%
  pivot_longer(temp, tss, names_to = "parameter")
```

```{r}
ggplot(sediment_temp_df) +
  geom_line(aes(x = sample_date_ymd, 
                y = temp),
            color = "red") + 
  geom_line(aes(x = sample_date_ymd,
                y = tss),
            color = "blue")
```

Over time how to total sediment and carbon vary and is there a correlation between the two.

<https://www.nature.com/articles/s41598-018-33640-3>

https://link.springer.com/article/10.1007/s40996-020-00460-3

# Different Method

```{r}

sediment_temp_df <- stream_water_df %>%
  select(-c(gage_ht, p_h, cond, cl, no3_n,
            so4_s, na, k, mg, ca, nh4_n, po4_p,
            tdn, tdp, si_o2, don, dic, doc,
            chloride_code, nitrate_code, 
            sulfate_code, sodium_code, 
            potassium_code, magnesium_code, calcium_code,
            nh4code, po4code, doc_code,
            dic_code, tdn_code, si_o2code,
            don_code)) %>%
  mutate(tss = as.numeric(tss)) %>% mutate(year = year(sample_date)) 
#made a column with year

sediment_temp_long <- sediment_temp_df %>%
  pivot_longer(cols = c(temp, tss), names_to = "parameter") #put into long format
%>%
  group_by(year,parameter) %>% 
  summarize( mean_value = mean(value, na.rm = TRUE)) # take average of temp and tss
```

```{r}

ggplot(sediment_temp_long, aes(x = year, y = mean_value)) + geom_line() +
  facet_wrap(~parameter, scales = "free", ncol = 1)
```
