---
title: "Quarto_Key"
format: html
editor: visual
---

# Loading Libraries

```{r}
library(tidyverse)
library(janitor)

```

# Documentation

Group:

1\) Created shared folder on Taylor-Adelaide

2\) Created github repository and shared with collaborators-Ale

3)Created Quarto graph with sections to work in separately

Andre - work on graph

# 

# Importing Data

```{r}

#all Quebrada
setwd("/courses/EDS214/pr/data/knb_lter_luq_raw")
p_r_m <-read_csv("RioMameyesPuenteRoto.csv", na = c("NA")) %>% clean_names()
b_q_1 <- read_csv("QuebradaCuenca1-Bisley.csv", na = c("NA")) %>% clean_names()  
b_q_2 <- read_csv("QuebradaCuenca2-Bisley.csv", na = c("NA")) %>% clean_names()
b_q_3 <-  read_csv("QuebradaCuenca3-Bisley.csv", na = c("NA")) %>% clean_names()

#Full Join

stream_water_df <- p_r_m %>% 
  full_join(b_q_1) %>% 
  full_join(b_q_2) %>%
  full_join(b_q_3)

setwd("/courses/EDS214/pr/data/processed_data")
write_csv(stream_water_df, "stream_water.csv")
```

# Tidying Data

```{r}

```

# Export Data

```{r}

```

# Pulling in Merged Data

```{r}

```

# Recreate Graph (Step 1)

## WORK UNDER HERE

```{r}

```

## WORK UNDER HERE ALSO

```{r}

```

# New Question

## WORK UNDER HERE ALE

```{r}

```

## WORK UNDER HERE ADELAIDE

```{r}

```

Over time how to total sediment and carbon vary and is there a correlation between the two.

<https://www.nature.com/articles/s41598-018-33640-3>

# 
